---
title: "Blog analytics and visuals"
author: "Christopher Kimani"
date: "2022-07-22"
output: word_document
  word_document: default
  html_document: default
  pdf_document: default
---
## 1. Defining the Question**
a) Specifying the Data Analytic Question

A Kenyan entrepreneur has created an online cryptography course and 
would want to advertise it on her blog. She currently targets audiences 
originating from various countries. In the past, she ran ads to advertise a 
related course on the same blog and collected data in the process. She 
would now like to employ your services as a Data Science Consultant to 
help her identify which individuals are most likely to click on her ads.

**b) Defining the Metric for Success**

1.Define the question, the metric for success, the context, experimental design taken and the appropriateness of the available data to answer the given question.

2.Perform univariate analysis.

3.Exhaustively perform bivariate analysis.

**c) Understanding the context**
Perform Exploratory Data Analysis for the give data set 
<http://bit.ly/IPAdvertisingData>

**d) Experimental design taken**

1.Reading and checking our data
2.Clean data by finding and dealing with outliers, anomalies, and missing 
data within the dataset.
3.Perform univariate and bivariate analysis.
4.From the insights provide a conclusion and recommendation.

**e) Appropriateness of the data**

The data is relevant for this study.

## 2. Loading the data**
```{r}
data <- read.csv('C:\\Users\\USER\\Downloads\\advertising.csv', header = TRUE)
```

checking the first 6 rows 
```{r}
head(data)
```
checking the last 6 rows 
```{r}
tail(data)
```
Checking the dimensions of the data
There are 1000 rows and 10 columns.
```{r}
dim(data)
```
Checking the column names 
```{r}
ls(data)
```
Checking the data types and structure 
The data is made up of numericals, characters and integers.
```{r}
str(data)
```
Checking the class of the data 
the data is a data frame
```{r}
class(data)
```
## 3. Data Cleaning
Checking for missing values
**There are no missing values in this data** 
```{r}
colSums(is.na(data))
```
Checking for duplicated values 
**There are no duplicated values in the data.**
```{r}
sum(duplicated(data))
```
Checking for outliers 
**There are no outliers in all other columns other than in the Area Income column.**
**I will however not remove these outliers as they may be useful in the analysis**
```{r}
length(boxplot.stats(data$'Daily.Time.Spent.on.Site')$out)
length(boxplot.stats(data$'Age')$out)
length(boxplot.stats(data$'Area.Income')$out)
length(boxplot.stats(data$'Daily.Internet.Usage')$out)
```
Plotting outliers in the Area Income column 
```{r}
boxplot(data$'Area.Income')
```
# Exploratory Data Analysis
## 4. Univariate Analysis
A summary of the data
**This shows a summary statistic of data. The mean, median, minimum and maximum values, the class of the data and the quartiles**
```{r}
summary(data)
```
Obtaining the variances of the data 
**This shows the variances of the chosen columns**
```{r}
var(data$'Age')
var(data$'Daily.Time.Spent.on.Site')
var(data$'Area.Income')
var(data$'Daily.Internet.Usage')
```
Obtaining the standard deviation of the data
**This shows the standard deviation of the data**
```{r}
sd(data$'Age')
sd(data$'Daily.Time.Spent.on.Site')
sd(data$'Area.Income')
sd(data$'Daily.Internet.Usage')
```
Obtaining the mode of the data 
**This shows the mode of the data**
```{r}
getmode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]}
getmode(data$'Age')
getmode(data$'Daily.Time.Spent.on.Site')
getmode(data$'Area.Income')
getmode(data$'Daily.Internet.Usage')
```
Getting which countries had the most count

**Aruba had the least count**
**France and Czech Republic had the most count**
```{r}
country = table(data$Country)
countries <- sort(country, increasing = TRUE)
countries <- sort(country, decreasing = TRUE)
head(countries)
```




Obtaining the count and visualisation of Click.on.Ad
**There is an equal number of those who clicked and those not clicked of 500**
**There is no class imbalance**
```{r}
clicked <- table(data$'Clicked.on.Ad')
clicked
labels <- c('clicked','did not click')
barplot(clicked, ylab = 'count', names.arg = labels, xlab = 'clicked', main ='clicked_count', col = 'green')
```
Gender count
**The number of females is slightly more than that of men**
```{r}
gender <- table(data$'Male')
gender
label <- c('female','male')
barplot(gender, ylab = 'count',names.arg = label, xlab = 'gender', main = 'gender_count', col = 'green')
```
Histogram for Age
**Most people in the data are aged 25-45 years**
```{r}
hist(data$'Age', ylab = 'count', xlab = 'age', main = 'age_count', col = 'green')
```
Histogram for Area Income 
**This chart shows that Area.Income is skewed to the right**
```{r}
hist(data$'Area.Income', ylab = 'count', xlab = 'income', main = 'area_income', col = 'green')
```
Showing the skewness of the Area Income column
```{r}
install.packages("moments")
library(moments)
skewness(data$Area.Income)
```
Histogram of Daily Internet Usage
**The data is more or less equally distributed**
```{r}
hist(data$'Daily.Internet.Usage', ylab ='count', xlab = 'internet usage', main = 'daily internet usage', col = 'green')
```
Histogram of Daily Time Spent on Site 
**The data is fairly skewed to the right**
```{r}
hist(data$Daily.Time.Spent.on.Site, ylab = 'count', xlab = 'daily time spent on site', main = 'Daily time spent on site', col = 'green')
skewness(data$'Daily.Time.Spent.on.Site')
```
##5. Bivariate Analysis

Average Income per gender 

**Men had more income than females**
```{r}
mean(data[data$Male == '1', 'Area.Income'])
mean(data[data$Male == '0', 'Area.Income'])
```

Scaling some columns 
```{r}
nage <- scale(age)
nusage <- scale(usage)
ntime <- scale(time)
```

Relationship between age and daily time spent on site 

**From this plot it is seen that people around the age of 25-35 years spent the most time on site**
**People above the age of 40 spend lesser time on site.**
**There is a negative correlation between age and time.**
```{r}
age <- data$'Age'
time <- data$'Daily.Time.Spent.on.Site'
plot(age, time, ylab = 'time spent', xlab = 'age', main = 'time spent on site against age', col = 'green')
cov(nage, ntime)
```

Relationship between age and daily internet usage

**The ages below 40 had more usage than those above 40 years.**
**There is a negative correlation between age and daily internet usage**

```{r}
usage <- data$'Daily.Internet.Usage'
cov(usage, age)
plot(age, usage, ylab ='daily internet usage', xlab ='age', main ='daily internet usage against age', col = 'green')
cov(nage, nusage)
```
# 8. Conclusion

1.  There are more females than males in our data.

2.  500 people clicked on the ads while 500 others did not click on the ads.

3.  The average area income is 55000.

5.  The average age of most audience is 36 years 

6.  Lisamouth and Williamsport cities both had the highest number of individuals in the dataset.



## 9. Recommendations

1.  Persons aged between 25 and 35 years old were the most in the data, thus creating ads to target these age group would be very impactful.

2.  Creating ads that target men makes more sense since men have more income compared to women.













































































































































































































































































